---
title: "proteoSE QC report"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  html_document:
    self_contained: true
    theme: readable
    toc: true
    toc_float: true
params:
  se: NULL
  group: "condition"
  batch: NULL
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      results = "asis", fig.width = 8, fig.height = 5,
                      fig.align = "center")

se       <- params$se
group    <- params$group
batch    <- params$batch
indicate <- intersect(c(group, batch, "replicate"),
                      colnames(SummarizedExperiment::colData(se)))

# Degrade, don't fail. A panel whose prerequisite is missing says so in one
# line and the report carries on. ggplots are returned and must be printed;
# the ComplexHeatmap/gridExtra panels (plot_missval, plot_detect, plot_cor)
# draw themselves and must not be printed twice.
panel <- function(expr) {
  tryCatch({
    p <- expr
    if (inherits(p, "ggplot")) print(p)
  }, error = function(e) {
    cat("\n*Not available:* ", conditionMessage(e), "\n\n", sep = "")
  })
  invisible(NULL)
}
```

## Overview

```{r overview}
mat <- SummarizedExperiment::assay(se)
cat("| | |\n|---|---|\n",
    "| Features | ", nrow(se), " |\n",
    "| Samples | ", ncol(se), " |\n",
    "| Assays | ", paste(SummarizedExperiment::assayNames(se), collapse = ", "), " |\n",
    "| Reported assay | ", SummarizedExperiment::assayNames(se)[1], " |\n",
    "| Missing values | ", sprintf("%.1f%%", 100 * mean(is.na(mat))), " |\n",
    "| Complete features | ", sum(stats::complete.cases(mat)), " |\n",
    sep = "")
```

### Design

```{r design}
print(knitr::kable(as.data.frame(SummarizedExperiment::colData(se))))
```

## Identifications

How many features each run saw. A sample well below its neighbours is a run
that underperformed, not a biological result.

```{r numbers}
panel(DEP2::plot_numbers(se))
```

How complete the matrix is: features binned by the number of samples they were
detected in.

```{r frequency}
panel(DEP2::plot_frequency(se))
```

```{r coverage}
panel(DEP2::plot_coverage(se))
```

## Missing values

Where the missingness sits, and whether it looks random. Black is a measured
value, white a missing one.

```{r missval, fig.height = 6}
panel(DEP2::plot_missval(se))
```

Intensity of fully observed features against those with missing values. If the
features with gaps are systematically the low-intensity ones, the missingness is
MNAR and imputation must account for that.

```{r detect, fig.height = 7}
panel(DEP2::plot_detect(se))
```

```{r randna}
randna <- tryCatch(
  SummarizedExperiment::rowData(proteoSE::add_randna(se))$randna,
  error = function(e) NULL
)
if (is.null(randna)) {
  cat("*MAR/MNAR classification not available.*\n")
} else {
  cat(sprintf(
    "`add_randna()` calls %d of %d features missing at random (MAR) and %d missing not at random (MNAR).\n",
    sum(randna), length(randna), sum(!randna)))
}
```

## Intensity distributions

Per-sample distributions. These should be comparable across samples before any
differential test; if they are not, normalise first.

```{r normalization, fig.height = 6}
panel(DEP2::plot_normalization(se))
```

## Replicate variability

Coefficient of variation within each condition. A condition with a visibly
wider spread has a replicate problem.

```{r cvs, fig.height = 6}
panel(DEP2::plot_cvs(se))
```

## Sample relationships

Do the samples group the way the design says they should? The correlation
panel needs a complete matrix -- on a raw assay it will say so; re-run with
`assay = ` set to an imputed assay to see it.

```{r pca}
panel(DEP2::plot_pca(se, indicate = indicate, n = min(500, nrow(se))))
```

```{r cor, fig.height = 6}
panel(DEP2::plot_cor(se, significant = FALSE, indicate = indicate))
```

---

```{r session}
cat("<details><summary>Session info</summary>\n\n```\n")
print(utils::sessionInfo())
cat("\n```\n</details>\n")
```
